---
title: "LFSC_521_R assignment"
author: "Yudan Shi"
date: "2022/2/6"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

* I'll use data from the **nycflights13** package, which is the dataset on flights departing New York City in 2013.

* Analyze the data using the **tidyverse** package, involving *split-apply-combine approach*, and using **ggplot2** to visualize the data.

* Since flight delays usually have a chain reaction, i.e., even if the initial problem that caused the delay has been resolved, the already delayed preceding flight could cause the following flight to be delayed as well. 

* Therefore, I will analyze the relationship between the delay time of previous fights and of the subsequent flights.

## Install and Library

```{r Install and library, message=FALSE, warning=FALSE}
# install.packages("tidyverse")
# install.packages("nycflights13")

library(tidyverse)
library(nycflights13) # Import the dataset
library(ggplot2)
```


## Analysis and Results

Firstly, I'll calculate the departure delay of the preceding flight from the same airport using `lag()` function based on the same *origin*.

```{r}
(lagged_delays <- flights %>% 
  arrange(origin, month, day, dep_time) %>%
  group_by(origin) %>%
  mutate(dep_delay_lag = lag(dep_delay)) %>%
  filter(!is.na(dep_delay),!is.na(dep_delay_lag))) ## remove rows including NA values
```

Then, grouped by delay time of the preceding flight, I'll calculate the average delay time of the subsequent flights. 

And then make a plot to visualize the relationship.

```{r}
lagged_delays %>%
  group_by(dep_delay_lag) %>%
  summarise(dep_delay_mean = mean(dep_delay)) %>%
  ggplot(aes(y = dep_delay_mean, x = dep_delay_lag)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 1500, by = 120)) +
  labs(y = "Departure Delay (min)", x = "Previous Departure Delay (min)") + 
  theme_classic()
```

## Discussion

The plot shows the relationship between the mean delay of a flight and delay of the preceding flight. 

For delays less than two hours, the relationship between the delay of the preceding flight and the current flight is nearly a line. 

After that the relationship becomes more variable, as long-delayed flights are interspersed with flights leaving on-time. 

After about 8-hours, a delayed flight is likely to be followed by a flight leaving on time.
